---
title: "HW 4"
author: "Gaea Daniel"
date: "10/18/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tableone)
library(Matrix)
library(survey)
library(survival)
library(testthat)
library(dummies)
library(grid)
library(Matching)
library(reshape2)
library(ggplot2)
library(knitr)
library(geepack)
library(lme4)
library(tidyverse)
library(haven)
library(lattice)

#Load data
help.spss <- read_csv("helpmkh.csv")
data(help.spss)
```

```{r}
#To access R-commander
library(Rcmdr)
library(splines)
library(RcmdrMisc)
library(car)
```

### Perform a Simple Linear Regression for:
OUTCOME variable indtot: "Inventory of Drug Use Consequences (InDue) total score - Baseline"
PREDICTOR variable mcs: "SF36 Mental Composite Score - Baseline"

```{r}
#Simple linear regression
RegModel.1 <- lm(indtot~mcs, data=helpmkh)
summary(RegModel.1)
```

Decide if you want to transform either variable indtot or mcs and if so, what transformation you applied and why.

Histograms were used to check the distribution of indtot and mcs. Mcs had a fairly normal distribution, but indtot was clearly left-skewed. Transformations (log, natural log, the mathematical constant, and square root) of indtot were created to normalize the distribution.
```{r}
#Histogram for indtot to check distribution
with(helpmkh, Hist(indtot, scale="frequency", breaks="Sturges", 
  col="darkgray"))

#Histogram for mcs to check for distribution
with(helpmkh, Hist(mcs, scale="frequency", breaks="Sturges", 
  col="darkgray"))

#Log transformation and histogram for indtot
helpmkh$indtotlog <- with(helpmkh, log10(indtot))
with(helpmkh, Hist(indtotlog, scale="frequency", 
  breaks="Sturges", col="darkgray"))

#Natural log transformation and histogram for indtot
helpmkh$indtotlog <- with(helpmkh, log(indtot))
with(helpmkh, Hist(indtotlog, scale="frequency", 
  breaks="Sturges", col="darkgray"))

#e transformation and histogram for indtot
helpmkh$indtote <- with(helpmkh, exp(indtot))
with(helpmkh, Hist(indtote, scale="frequency", breaks="Sturges", 
  col="darkgray"))

#Square root transformation and histogram for indtot
helpmkh$indtotsqrt <- with(helpmkh, sqrt(indtot))
with(helpmkh, Hist(indtotsqrt, scale="frequency", 
  breaks="Sturges", col="darkgray"))
```

The square root transformation of indtot appeared to normalize the indtot variable the best and was used for the simple linear regression.

```{r}
#Simple linear regression with square root transformation of indtot and mcs
RegModel.5 <- lm(indtotsqrt~mcs, data=helpmkh)
summary(RegModel.5)
```

### Perform regression diagnostics:
Check the normality of the residuals (histogram and Q-Q plots)

```{r}
#Histogram and Q-Q plots for indtot
with(helpmkh, Hist(indtot, scale="frequency", breaks="Sturges", 
  col="darkgray"))
with(helpmkh, qqPlot(indtot, dist="norm", id.method="y", id.n=2, 
  labels=rownames(helpmkh)))

#Histogram and Q-Q plots for mcs
with(helpmkh, Hist(mcs, scale="frequency", breaks="Sturges", 
  col="darkgray"))
with(helpmkh, qqPlot(mcs, dist="norm", id.method="y", id.n=2, 
  labels=rownames(helpmkh)))
```

Check for linearity - is there any systematic relationship between the residuals and the predicted (or fitted) values? No

```{r}
#Whitout the line
scatterplot(indtot~mcs, reg.line=FALSE, smooth=FALSE, 
  spread=FALSE, boxplots=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), data=helpmkh)

#With the line
scatterplot(indtot~mcs, reg.line=lm, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  data=helpmkh)
```

Homoscedasticity - plot of standardized residuals versus fitted values - this is known as a "Scale-Location" graph. Check for outliers and data points with high leverage or influence: outliers are often identified with standardized residuals > 3 (or <-3) and influential observations are often identified using Cook's D

```{r}
# Change the layout to 2x2 to accommodate all plots
par(mfrow=c(2,2))
par(mar = rep(2, 4))
 
# Diagnostic Plots
plot(RegModel.5)
```

    Provide a summary of the regression results.
        provide a FIGURE of the model, in this case a scatterplot with the fitted line overlaid and 95% confidence intervals if you can
        Make a TABLE presenting the fitted regression model (coefficients and tests of significance for those coefficients)
        describe the variance explained by the model (based on r2)
        describe the model itself based on the y-intercept and slope terms
        note any limitations or issues with the model fit or interpretation of the model
See above

    Perform a One-way ANOVA for:
        OUTCOME variable indtot: "Inventory of Drug Use Consequences (InDue) total score - Baseline"
        GROUP variable racegrp: "Racial Group of Respondent"
        I would suggest merging "other" and "hispanic" together and create a 3-group variable for race, since the "other" category is only about 6% of the sample.
        options - you can use either an ANOVA or GLM modeling approach
        if the GROUP variable is significant, also perform post hoc tests - use some kind of pairwise error rate adjustment (i.e. bonferroni, sidak, Tukey's HSD, etc) - be sure to report which one you used and why
Bonferroni is used because the significant difference may not exist between all conditions.
        
```{r}
#One-way ANOVA
AnovaModel.7 <- aov(indtot ~ racegrp, data=helpmkh)
summary(AnovaModel.7)
with(helpmkh, numSummary(indtot, groups=racegrp, 
  statistics=c("mean", "sd")))

#Bonferroni
outlierTest(AnovaModel.7)
```

    Perform model diagnostics:
        homoscedasticity - look at a test for equal variance (Levene's test or Bartlett's test).
        if this test of equal variances fails, you may want to report a modified F-test (e.g. Welch's test)
        Present a summary of the ANOVA results.

```{r}
#Bartlett's test
with(helpmkh, tapply(indtot, racegrp, var, na.rm=TRUE))
bartlett.test(indtot ~ racegrp, data=helpmkh)

#Levene's test
with(helpmkh, tapply(indtot, racegrp, var, na.rm=TRUE))
leveneTest(indtot ~ racegrp, data=helpmkh, center="median")

#Two-variance F-test
with(helpmkh, tapply(indtot, treat,  var, na.rm=TRUE))
var.test(indtot ~ treat, alternative='two.sided', conf.level=.95,
   data=helpmkh)
```
    
        Make a FIGURE of the group mean differences - either an error-bar plot or a series of boxplots one for each group to show the group differences in the outcome
        
```{r}
#Boxplot
with(helpmkh, tapply(indtot, treat, var, na.rm=TRUE))
leveneTest(indtot ~ treat, data=helpmkh, center="median")
Boxplot(indtot~racegrp, data=helpmkh, id.method="y")
```

        Make a TABLE presenting the ANOVA results
        describe the model results - was the GROUP (racegrp) significant? Yes
        If GROUP is significant, what did the post hoc tests reveal? The post hoc test (Bonferonni) revealed that it was not.